On Extensibility and Personalizability of Multi-Modal Trip Planning

### 2016-03-02 paper.tex
Small save.
Updated section Preferences.

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tblref}[1]{Table~\ref{tbl:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\thmref}[1]{Theorem~\ref{thm:#1}}
\newcommand{\propref}[1]{Proposition~\ref{prop:#1}}
\newcommand{\corref}[1]{Corollary~\ref{cor:#1}}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{On Extensibility and Personalizability of Multi-Modal Trip Planning}
\author{
	Xudong Liu\\
	Department of Computer Science\\
	University of Kentucky\\
	329 Rose Street\\
	Lexington, KY 40506\\
	USA
	\And
	PARC People\\
	Palo Alto Research Center\\
	3333 Coyote Hill Road\\
	Palo Alto, CA 94304\\
	USA
}
\maketitle
\begin{abstract}
	Trip planning is a practical task that has drawn
	extensive attention from the AI planning and scheduling community
	and the industrial/commercial sectors.
	In this paper, we consider the setting of the \tit{multi-modal}
	trip planning, where users can exploit different transportation
	modes, such as walking, biking, public transit, and taxi.
	In such a context, it would be of benefit if the
	user was able to extend the \tit{cost base}, including traveling time 
	and fare, of the planner, and to personalize the planner according
	to her own constraints and preferences.
	To this end, we designed and developed a hipergraph-based multi-modal
	trip planner that allows users to upload auxiliary cost metrics (e.g.,
	crime rates), to specify constraints as a theory in the 
	\tit{linear temporal logic}, and to express preferences as a 
	\tit{preferential cost function}.
\end{abstract}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/result1.png}
  \caption{State transition diagram\label{fig:trjs}}
\end{figure}

@@

A state is described as a set of \tit{state variables}.
The state variables of a state $S$ include the transportation mode $M$ that led to $S$,
time spent $T$ so far per mode (e.g., $T_\bike$ for biking and $T_\public$ for
public transit), fare $D$ spent so far per mode (e.g., $D_\gas$ for driving and
$D_\taxi$ for taking a cab), and variables related to the auxiliary data once uploaded.
These extra data related variables are metrics such as the sum ($A_\SUM$),
the maximum ($A_\MAX)$, the minimum ($A_\MIN$), and the average ($A_\AVG$) data along the path.

We focused on weighted functions over state variables and
designed the cost function, called preferential cost function (PCF), that guides the
graph-based search engine in our trip planner as follows.
\begin{equation}
	\begin{aligned}
		\PCF(S) =& \beta_1 * (\alpha_1 \cdot T_\walk + \alpha_2 \cdot T_\wait + \ldots) \\
								&+ (D_\gas + D_\public + \ldots) \\
								&+ \beta_2 * (A_\SUM + \ldots),
	\end{aligned}
\end{equation}
where $\alpha_i$ are real numbers representing the relations among different time pieces,
and $\beta_1$ ($\beta_2$) is the ratio that essentially describes how much in dollars a user would pay to
save an hour (an auxiliary data, respectively).

\subsubsection{Preference Elicitation}
To gather these coefficients ($\alpha_i$ and $\beta_i$) in our $\PCF$, we designed interface to
elicit these numbers from the user.

>>>>

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tblref}[1]{Table~\ref{tbl:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\thmref}[1]{Theorem~\ref{thm:#1}}
\newcommand{\propref}[1]{Proposition~\ref{prop:#1}}
\newcommand{\corref}[1]{Corollary~\ref{cor:#1}}
\newcommand{\eqtref}[1]{Equation~\ref{eq:#1}}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{On Extensibility and Personalizability of Multi-Modal Trip Planning}
%\author{
%	Xudong Liu\\
%	Department of Computer Science\\
%	University of Kentucky\\
%	329 Rose Street\\
%	Lexington, KY 40506\\
%	USA
%	\And
%	PARC People\\
%	Palo Alto Research Center\\
%	3333 Coyote Hill Road\\
%	Palo Alto, CA 94304\\
%	USA
%}
\maketitle
\begin{abstract}
	Trip planning is a practical task that has drawn
	extensive attention from the AI planning and scheduling community
	and the industrial/commercial sectors.
	In this paper, we consider the setting of the \tit{multi-modal}
	trip planning, where users can exploit different transportation
	modes, such as walking, biking, public transit, and taxi.
	In such a context, it would be of benefit if the
	user was able to extend the \tit{cost base}, including traveling time 
	and fare, of the planner, and to personalize the planner according
	to her own constraints and preferences.
	To this end, we designed and developed a hipergraph-based multi-modal
	trip planner that allows users to upload auxiliary cost metrics (e.g.,
	crime rates), to specify constraints as a theory in the 
	\tit{linear temporal logic}, and to express preferences as a 
	\tit{preferential cost function}.
\end{abstract}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.4\textwidth]{figs/trajectories.pdf}
  \caption{State transition diagram\label{fig:trjs}}
\end{figure}

@@

A state is described as a set of \tit{state variables}.
The state variables of a state $S$ include the transportation mode $M$ that led to $S$,
time $T$ spent so far per mode (e.g., $T_\bike$ for biking and $T_\public$ for
public transit), fare $D$ spent so far per mode (e.g., $D_\gas$ for driving and
$D_\taxi$ for taking a cab), and variables related to the auxiliary data once uploaded.
These extra data related variables are metrics such as the sum ($A_\SUM$),
the maximum ($A_\MAX)$, the minimum ($A_\MIN$), and the average ($A_\AVG$) data along the path.

We focused on weighted functions over state variables and
designed the cost function, called preferential cost function (PCF), that guides the
graph-based search engine in our trip planner as follows.
\begin{equation}
	\begin{aligned}
		\PCF(S) =& \beta_1 * (\alpha_\walk \cdot T_\walk + \alpha_\bike \cdot T_\bike + \ldots) \\
								&+ (D_\gas + D_\public + \ldots) \\
								&+ \beta_2 * (A_\SUM + \ldots),
	\end{aligned}
	\label{eq:pcf}
\end{equation}
where $\alpha_i$ are real numbers representing the relations among different time pieces,
and $\beta_1$ ($\beta_2$) is the ratio that essentially describes how much in dollars a user would pay to
save an hour (an auxiliary data, respectively).

\subsubsection{Preference Elicitation}
To gather these coefficients ($\alpha_i$'s and $\beta_i$'s) in our $\PCF$, we designed interface to
elicit these numbers from the user.
It asks the user questions and collect answers from the user to derive the coefficients.
These questions are as follows.
\begin{enumerate}
	\item One hr of walking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of biking $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of public transit $=$ \underline{\hspace{1cm}} hrs of driving?
	\item One hr of taxi $=$ \underline{\hspace{1cm}} hrs of driving?
	\item How much in dollars would you pay to save an hour in traveling?
	\item How much in dollars would you pay to avoid an auxiliary event (e.g., crime) in traveling?
\end{enumerate}
For instance, the user answers 3, 2, 0.25 and 0.5 to the first four questions in the list above.
Intuitively, the numbers indicate that the user prefers public transit the most, followed by
taxi, driving, biking and walking, in order.
We show how we can now derive $\alpha_i$'s in \eqtref{pcf}.
We start with setting $\alpha_\car=1$.
Now, since one hour of walking is equivalent to 3 hours of walking,
we have $\alpha_\walk \times 1 = \alpha_\car \times 3$;
hence, we derive $\alpha_\walk = 3$.
Similarly, we have $\alpha_\bike = 2$, $\alpha_\public = 0.25$,
and $\alpha_\taxi = 0.5$.
Computing $\beta_1$ and $\beta_1$ is straightforward.
Indeed, function $\PCF$ with the input of time, fare and
auxiliary metric pieces boils down to monetary cost,
and the planner computes the best path by optimize
based on this overall monetary cost in the searching
process.
